Assessing landscape N removal in coastal New England catchments using the

### 2023-09-07 manuscript/manuscript.Rmd
submission ready!  cache on, sessioninfo, akncowledgements, and code statement.

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  cache = FALSE,
  tidy = FALSE,
  message = FALSE,
  fig.height = 4.8,
  fig.width = 6
)
#load(system.file("testdata.rda", package = "nsink"))
#load("testdata.rda")
```

@@

The `nsink` R package was written to simplify the acquisition and management of the spatial data necessary to estimate N removal within each identified landscape sink, estimate cumulative N removal along a specified flowpath, and create a set of static watershed maps [@hollister2022nsink] . Datasets used by `nsink` include the National Hydrography Dataset Plus (NHDPlus), Soil Survey Geographic Database (SSURGO), the National Land Cover Dataset (NLCD) land cover and the National Land Cover Dataset (NLCD) impervious surface [@jin2019overall; @moore2019user; @soil2017web]. The static maps illustrate 1) N Removal Efficiency, a color-coded map of the estimated N removal efficiency of the different landscape sinks; 2) N Transport Index, a heat map with the cumulative N removal along flowpaths originating from sources represented by a grid of points across a watershed. This heat map highlights “leaky” source areas with less downstream N retention versus those with higher downstream retention; and 3) N Delivery Index, combining maps (1) and (2) along with source loading based on NLCD categories and expressed as an index ranging from 0 to 1, resulting in a map showing potential N delivery from different sources, after accounting for the potential N removal as water moves downstream. N-Sink and `nsink` integrate these data to make the application of complex hydrologic and biogeochemical processes accessible to end users and stakeholders in a visually clear and useful format. Here we describe the development of the package, and example workflow, and simple use applications to demonstrate the capabilities of the tools.

@@

>>>>

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  tidy = FALSE,
  message = FALSE,
  fig.height = 4.8,
  fig.width = 6
)
#load(system.file("testdata.rda", package = "nsink"))
#load("testdata.rda")
```

@@

The `nsink` R package was written to simplify the acquisition and management of the spatial data necessary to estimate N removal within each identified landscape sink, estimate cumulative N removal along a specified flowpath, and create a set of static watershed maps [@hollister2022nsink] . Datasets used by `nsink` include the National Hydrography Dataset Plus (NHDPlus), Soil Survey Geographic Database (SSURGO), the National Land Cover Dataset (NLCD) land cover and the National Land Cover Dataset (NLCD) impervious surface [@jin2019overall; @moore2019user; @soil2017web]. The static maps illustrate 1) N Removal Efficiency, a color-coded map of the estimated N removal efficiency of the different landscape sinks; 2) N Transport Index, a heat map with the cumulative N removal along flowpaths originating from sources represented by a grid of points across a watershed. This heat map highlights “leaky” source areas with less downstream N retention versus those with higher downstream retention; and 3) N Delivery Index, combining maps (1) and (2) along with source loading based on NLCD categories and expressed as an index ranging from 0 to 1, resulting in a map showing potential N delivery from different sources, after accounting for the potential N removal as water moves downstream. N-Sink and `nsink` integrate these data to make the application of complex hydrologic and biogeochemical processes accessible to end users and stakeholders in a visually clear and useful format. Here we describe the development of the package, and example workflow, and simple use applications to demonstrate the capabilities of the tools. The code for this analysis is available from <https://github.com/USEPA/nsink_manuscript/> and is archived at <https://doi.org/10.5281/zenodo.8325360>.  All analyses were conducted with R version 4.2.2 and details on R package versions and operating system used for this analysis are included in a file, `sessioninfo.txt` at <https://github.com/USEPA/ri_wq_trends/blob/master/sessioninfo.txt> [@r2022; @here; @ggplot2; @sf1; @sf2; @knitr1; @knitr2; @knitr3; @feddata; @hollister2022nsink; @units; @raster; @terra; @tidyterra; @lubridate; @units].

@@

```{r session_info, echo=FALSE}
#Send session info to a file: 
capture.output(sessionInfo(),file=here::here("sessioninfo.txt"))
```

Acknowledgements
================

Many people have contributed in various ways to the development of the N-Sink 
concept. In particular, we would like to thank Peter August and Chris Damon of 
the University of Rhode Island’s Department of Natural Resources Science.  
Additionally, we are grateful to Stephen Shivers, Anne Kuhn, Andrea Bartolotti, 
Suzanne Ayvazian, and Tim Gleason for constructive reviews of this paper. The 
views expressed in this article are those of the authors and do not necessarily 
represent the views or policies of the U.S. Environmental Protection Agency. Any 
mention of trade names, products, or services does not imply an endorsement by 
the U.S. Government or the U.S. Environmental Protection Agency. The EPA does 
not endorse any commercial products, services, or enterprises. This contribution 
is identified by the tracking number ORD-054342 of the Atlantic Coastal 
Environmental Sciences Division, Office of Research and Development, Center for 
Environmental Measurement and Modeling, US Environmental Protection Agency.
